### 2018-03-16 PAPER/paper.tex

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused. Also, the teaser figure should only have the
%% width of the abstract as the template enforces it.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{gfx/}{./}} % where to search for the images

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
\usepackage{makecell}
\usepackage{multirow}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}
%% please declare the paper type of your paper to help reviewers, only shown in review mode
%% choices:
%% * algorithm/technique
%% * application/design study
%% * evaluation
%% * system
%% * theory/model
\vgtcpapertype{theory/model}

%% Paper title.
\title{Computing Graphical Perception}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
\authorfooter{
%% insert punctuation at end of each item
\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
E-mail: \{haehn,pfister\}@seas.harvard.edu.
%
\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
}

%other entries to be set up for journal
\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{TODO%
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Machine Perception, Deep Learning}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

%\CCScatlist{ % not used in journal version
% \CCScat{K.6.1}{Management of Computing and Information Systems}%
%{Project and People Management}{Life Cycle};
% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
%}

%% Uncomment below to include a teaser figure.
\teaser{
  \centering
  \includegraphics[width=\linewidth]{fish.jpg}
  \caption{Here is a fish.}
	\label{fig:teaser}
}

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

\vgtcinsertpkg

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command

\input{00_introduction.tex}

\input{01_related_work.tex}

\section{Experimental Setup}

The experiments shown in this paper are either regression or classification tasks. We formulate any estimation of quantities (e.g. angles, positions, lengths etc.) as a regression problem between $0$ and $1$. The output indicates the percentage in regards to the degrees of freedom of the individual experiment. If the experiment involves a choice, we formulate it as a classification problem.

\subsection{Measures}

\textbf{Accuracy.} We use the same metric as Cleveland McGill to measure accuracy.
\begin{equation}
	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
\end{equation}

\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding. For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.

\subsection{Classifiers}

Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network. We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation. These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
\\~\\
\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}). We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
\\~\\
\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP. We train the \emph{LeNet} classifier with tune it specifically towards each visualization. For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
\\~\\
\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}). We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.

\begin{table}[t]
\centering
\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task. This yields different sets of trainable parameters. We also train the MLP directly on the visualizations without any additional feature generation.}
\resizebox{\linewidth}{!}{
\begin{tabular}{lrl}
%	\toprule
%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
%	\midrule
%	MLP & $0$ & $2,560,513$ \\
%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
%	\bottomrule
	\toprule
	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
	\midrule
	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
	\bottomrule
\end{tabular}
}
\label{tab:parameters}
\vspace{-4mm}
\end{table}

\begin{figure}[t]
	\centering
	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU). We use Dropout regularization to prevent overfitting. We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables. The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv. layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv. layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv. layers, filter size $3\times3$) to increase the number of trainable parameters.}
	\label{fig:classifiers}
\end{figure}

\section{Elementary Perceptual Tasks}

THe Figure 1 of Cleveland McGill

\subsection{Parametrization}

Here we show each of Figure 1 with its parametrization.

\section{Position-Angle Experiment}

This is pie chart vs bar chart

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{fish.jpg}
  \caption{The Position-Angle Experiment}
	\label{fig:position_angle_experiment}
\end{figure}

\section{Position-Length Experiment}

This is the one where we estimate two selected bars compared to the longest one

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{fish.jpg}
  \caption{The Position-Length Experiment}
	\label{fig:position_length_experiment}
\end{figure}

\section{Bars and Framed Rectangles Experiment}

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{fish.jpg}
  \caption{The Bars and Framed Rectangles Experiment}
	\label{fig:bars_and_framed_rectangles_experiment}
\end{figure}

\subsection{Weber's Law}

\section{Results and Discussion}

\begin{figure*}[h]
	  \includegraphics[width=\linewidth]{figure1.pdf}
  \caption{Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}. We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
	\label{fig:figure1_results}
\end{figure*}

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
  \caption{Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts. We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
	\label{fig:position_angle_results}
\end{figure}

\begin{figure*}[t]
	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
  \caption{Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}. We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
	\label{fig:position_angle_results}
\end{figure*}

Cleveland McGills Ranking - can we observe something similar?

1. Position along a common scale e.g. scatter plot

2. Position on identical but nonaligned scales e.g. multiple scatter plots

3. Length e.g. bar chart

4.Angle \& Slope (tie) e.g. pie chart

5. Area e.g. bubbles

6. Volume, density, and color saturation (tie) e.g. heatmap

7. Color hue e.g. newsmap

\section{Conclusions}

Future work: allow insights for infovis for machines

%% if specified like this the section will be committed in review mode
%\acknowledgments{
%The authors wish to thank A, B, and C. This work was supported in part by
%a grant from XYZ (\# 12345-67890).}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{paper.bib}
\end{document}