### 2018-03-29 PAPER/02_experimental_setup.tex

\section{Experimental Setup}

We conduct quantitative experiments to measure the ``perception'' of low-level visual encodings such as positions, angles, curvature, and lengths for different convolutional neural networks. We treat each classifier configuration as a subject in a between-subjects experiment evaluate supervised learning algorithms. We formulate any estimation of quantities (e.g. angles, positions, lengths etc.) as a logistic regression problem with single or multiple outputs between $0$ and $1$. The output indicates the percentage in regards to the degrees of freedom of the individual experiment. As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).

\subsection{Measures}

\textbf{Accuracy.} Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
\begin{equation}
	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
\end{equation}
In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}). This allow a more direct comparison of percent errors.
\\~\\
\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding. For regression tasks the loss is defined as mean squared error (MSE).% and for classification tasks the loss is categorical cross-entropy.
\\~\\
\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.

\subsection{Classifiers}

Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network. We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation. These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
\\~\\
\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}). We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
\\~\\
\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP. We train the \emph{LeNet} classifier with tune it specifically towards each visualization. For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
\\~\\
\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}). We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
\\~\\
\noindent{\textbf{Environment.}} We run all experiments on an NVIDIA DGX1 machine with Tesla V100 graphical processing units. We use the KERAS framework with tensorflow.

\begin{table}[t]
\centering
\caption{\textbf{Classifier Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression. This yields different sets of trainable parameters. We also train the MLP directly on the visualizations without any additional feature generation.}
\resizebox{\linewidth}{!}{
\begin{tabular}{lrl}
%	\toprule
%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
%	\midrule
%	MLP & $0$ & $2,560,513$ \\
%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
%	\bottomrule
	\toprule
	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
	\midrule
	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
	\bottomrule
\end{tabular}
}
\label{tab:parameters}
\vspace{-4mm}
\end{table}
%\begin{figure}[t]
%	\centering
%	  \includegraphics[width=\linewidth]{classifiers.pdf}
%  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU). We use Dropout regularization to prevent overfitting. We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables. The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv. layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv. layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv. layers, filter size $3\times3$) to increase the number of trainable parameters.}
%	\label{fig:classifiers}
%\end{figure}
\begin{figure}[t]
	\centering

    \subfloat[Feature Generation]{
		\includegraphics[width=4.8cm,valign=c]{classifier_left.pdf}
		\vphantom{\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}}
	}
	\hfill
    \subfloat[Multilayer Perceptron]{
		\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}
	}

  \caption{\textbf{Classifier Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU). We use Dropout regularization to prevent overfitting. As output, we perform linear regression for continuous variables. The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv. layers, filter size $5\times5$), VGG19 (16 conv. layers, filter size $3\times3$), or Xception (36 conv. layers, filter size $3\times3$) to test different model complexities. }
	\label{fig:classifiers}
\end{figure}

\subsection{Data}

We create all visualizations as parametrized rasterized images without interpolation. The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}. 
\\~\\
\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
\\~\\
\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images. We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables. This eliminates leakage during training and evaluation. We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$. 
\\~\\
\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.