### 2018-03-30 PAPER/02_experimental_setup.tex
Section 2

\section{Experimental Setup}

We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths. We formulate these measurement tasks as logistic regression problems: given a stimuli image of an elementary visualization, the networks must estimate the single quantity present or the ratio between multiple quantities present.

For each experiment, we use a single factor between-subject design, with the factor being the network used. This lets us evaluate whether different network designs are competitive against existing human perception results. We train each network in a supervised fashion with a mean-squared error (MSE) loss between the ground-truth labels and the network's estimate of the measurement from observing the generated stimuli images. Then, we test each network's ability to generalize to new examples with a separate data, created using the same stimuli generator function but with unseen ground-truth measurements (Section~\ref{sec:data}).

% We define a series of hypotheses prior to each experiment.

\subsection{Networks}
\label{sec:networks}

\noindent{\textbf{Multilayer Perceptron.}} As baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}). Our MLP contains a layer of $256$ perceptrons, which are activated as rectified linear units (ReLU)~(Fig.~\ref{fig:classifiers}). We train this layer with dropout (probability $= 0.5$) to prevent overfitting, and then combine these ReLU units to regress our output measurement.
\\~\\
\noindent{\textbf{Convolutional Neural Networks.}} We compare different convolutional neural networks (CNNs) with both `trained from scratch' weights and pre-trained weights on a database of natural images (1000-class ImageNet \cite{imagenet}). These networks are the traditional LeNet-5 with 2 layers, which was designed to recognize hand-written digits~\cite{lenet}; the VGG19 network with 19 layers, which was designed to solve the ImageNet object recognition challenge~\cite{simonyan_very_deep2014}; and the Xception network with 36 layers~\cite{xception}, which was also designed to solve the ImageNet object recognition challenge plus the 15,000-class JFT object recognition challenge~\cite{Hinton2015}. Each of these networks has as its last layers an MLP architecture equivalent to our baseline, and so they act as earlier image and feature processors for this final regressor. Since the networks are of different architectures, the number of trainable parameters changes, with some networks having more capacity than others (Table~\ref{tab:parameters}).

For \emph{VGG19} and \emph{Xception}, we have two variants: the network trained from scratch on elementary perceptual tasks, plus the network using weights that were previously trained on the ImageNet object recognition challenge \emph{except} for the MLP layer. This is intended to produce early-layer features which mimic human vision, and then to see whether they are more or less useful than networks trained from scratch. 
\\~\\
\noindent{\textbf{Optimization.}} All network hyperparameters, optimization methods, and stopping conditions are fixed across networks (Table \ref{tab:parameters}). We train for $1000$ epochs using stochastic gradient descent with Nesterov momentum, but stop early if the loss does not decrease for ten epochs.
\\~\\
\noindent{\textbf{Environment.}} We run all experiments on Tesla X and Tesla V100 graphical processing units. We use the KERAS framework with a TensorFlow backend to train the networks, and use the scikit-image library to generate the stimuli. % and scikit-learn libraries - JT: What do we use this for?

\begin{table}[t]
\centering
\caption{\textbf{Network Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression. This results in different sets of trainable parameters. As a baseline, we also train the MLP directly on the visualization images without any additional feature generation.}
\resizebox{\linewidth}{!}{
\begin{tabular}{lrl}
%	\toprule
%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
%	\midrule
%	MLP & $0$ & $2,560,513$ \\
%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
%	\bottomrule
	\toprule
	Network & \makecell{Trainable\\Parameters} & Optimization \\
	\midrule
	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
	\bottomrule
\end{tabular}
}
\label{tab:parameters}
\vspace{-4mm}
\end{table}
%\begin{figure}[t]
%	\centering
%	  \includegraphics[width=\linewidth]{classifiers.pdf}
%  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU). We use Dropout regularization to prevent overfitting. We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables. The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv. layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv. layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv. layers, filter size $3\times3$) to increase the number of trainable parameters.}
%	\label{fig:classifiers}
%\end{figure}
\begin{figure}[t]
	\centering

    \subfloat[Feature Generation]{
		\includegraphics[width=4.8cm,valign=c]{classifier_left.pdf}
		\vphantom{\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}}
	}
	\hfill
    \subfloat[Multilayer Perceptron]{
		\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}
	}

  \caption{\textbf{Network Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU). We use Dropout regularization to prevent overfitting. As output, we perform linear regression for continuous variables. The MLP can learn to represent the visualizations directly, but we also learn features generated by LeNet (2 conv. layers, filter size $5\times5$), VGG19 (16 conv. layers, filter size $3\times3$), or Xception (36 conv. layers, filter size $3\times3$) to test different model complexities.}
	\label{fig:classifiers}
\end{figure}

\subsection{Data}
\label{sec:data}

\noindent\textbf{Image Stimuli and Labels} 
We create our stimuli visualizations as 100$\times$100 binary images, rasterized without interpolation. We write a parameterized stimuli generator for each elementary task. The number of possible parameter values differs per experiment, and we summarize these in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}. Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$. Then, we add subtle random noise (between $0--0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.

Each stimuli image also has an associated ground truth label representing the parameter set which generated the image, e.g., the length in pixels of a bar. As before, we scale these labels to the range of $0.0$ to $1.0$, which represent the maximum and minimum values that this parameter can take.
\noindent\textbf{Training/Validation/Test Splits.} For each task, we use 60,000 training images, 20,000 validation images, and 20,000 test images. To create these datasets, we generate stimuli from random parameters and add them to the sets until the target number is reached, while maintaining distinct (random) parameter spaces for each set to ensure that there is no leakage between training and validation/testing.

\subsection{Measures and Analysis}

\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments. We run every experiment seperately twelve times, and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test. 
%Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). -> JT: No it doesn't.
% Finally, we average the results over the runs. -> JT: Fine, but we'll do distributions anyway, so not worth saying.
\\~\\
\noindent{\textbf{Task Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy. To allow comparison between their human results and our machine results, we also use MLAE as a presentation metric:
\begin{equation}
	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
\end{equation}
In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}). This allow a more direct comparison of percent errors. However, please note that our networks were trained using MSE loss and not directly with MLAE.
\\~\\
\noindent{\textbf{Task Confidence Intervals.}} We follow Cleveland and McGill and present $95\%$ confidence intervals. However, as our networks but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity. JT: What? Why 97.5\% ? Why not 95\% ?
\\~\\
\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we analyze dependent variables using analysis of variance (ANOVA) followed by parametric tests. JT: Which tests?
\\~\\
\noindent{\textbf{Training Efficiency.}} We use the training convergence rate as a measure of how easy or hard a particular task is for the network to learn to solve. This is defined as the MSE loss decrease per training epoch, which is an indicator of the training efficiency of the network with respect to the visual encoding.
\\~\\
\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer quetsions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.

Further, some experiments compare different visual encoding types, e.g., bar plot vs.~stacked bar plot. We train and evaluate individual networks for each parameterization, plus we also train and evaluate a networks on stimuli across the different types. This single decision-making software better mimics the judgements that a human would be able to make. 
%This scenario affects the optimization process and result in networks with more flexible knowledge with the caveat of longer training times. -> JT: Save it for the results discussion.